The document title

### 2021-03-23 manuscript/manuscript.Rmd
Add FS suggestions

To assess the models prediction accuracy, mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE) were calculated. Also, methods agreement was evaluated with Bland-Altman plots by plotting the difference between the actual (FP measured) and predicted values against their mean. Bias was computed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the mean between actual and predicted values [@Bland1986]. In order to test whether there was a consistent over or underestimation of the acual values, one-sample \textit{t} tests were performed to assess if the difference between actual and predicted values are different than zero. Linear regressions were applied to identify if the magnitude of the mean between actual and predicted values influenced the magnitude of their difference. This was done to ascertain if the prediction accuracy was constant throughout the assayed magnitude range, which is known as proportional bias [@Giavarina2015]. These analyses were conducted separately for data from each accelerometer placement and outcome.

@@

Regression equations were developed to predict the mechanical loading variables based on accelerometry data. In the pGRF prediction equations, body mass and pACC were significant predictors and included in the models. Table 1 shows the regression coefficients, R^2^ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component. In all pGRF prediction models, R^2^ values were `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome varibale to a great extent. Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020]. RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N. Of all the pGRF predictions equations developed, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an R^2^ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).

@@

Also, to simplify the use of these equations and enhace their applicability, a software package was developed for the use in the R statistical software (REF). Included in this package are functionalities to read raw ACC data, process it and apply the equations to predict both pGRF and pLR. In the package website (ADDRESS) there are instructions on how to use it to obtain the predictions, namely in the 'Home' and 'Get started' sections.

@@

The purpose of this study was to develop pGRF and pLR prediction equations based on accelerometer data able to accurately estimate the mechanical loading during walking and running. The predictions were tested with accelerometers placed at different body locations and using both the resultant vector and its vertical component in a set of incremental speeds, ranging from slow walking to running. Our results demonstrate that accelerometer-based regression models are a valid and accurate method to estimate pGRF, especially with data from hip-worn accelerometers, during human locomotion. The results also indicate that pLR can also be predicted, however with a substantially lower accuracy compared with the pGRF models.

All pGRF models showed an accurate prediction, with a MAPE around 5.5\%, using data from hip-worn accelerometers, and around 6.0\% with data from the other placements. On the other hand, the pLR prediction models presented considerably higher errors, as shown by their MAPE around 29\%. Also, differently than the pGRF equations, the pLR equations that showed the best accuracy are from the lower back placement. Moreover, for both outcomes, predictions from the resultant vector tended to be slightly better than from its vertical component. These results are also reflected in the Bland-Altman plots, as the pGRF plots presented considerably narrower limits of agreement and lower dispersion compared to the pLR. Furthermore, it can be observed that the Bland-Altman plot fot the pGRF equation for the resultant vector and data from accelerometers placed at the hip presents a high degree of clustering around the zero and very few data points outside the limits of agreement, which confirms it to be the prediction equation that better performs.

In respect to the various accelerometer placements analysed in this study, the hip placement showed the best results, mainly in the pGRF models. This reinforces the hip placement as the established location to collect and analyse accelerometry-measured physical activity data, as this placement is already well accepted for energy expenditure prediction and physical activity intensity classification [@Migueles2017] and has shown a high mechanical loading prediction ability in this and previous studies [@Veras2020; @Neugebauer2018; @Neugebauer2012]. Besides the various placements, the current study also analysed predictions from both the resultant and vertical vectors. Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector, and, as can be observed in Supplemental Figure S1, values from both vectors in this study were very similar. Also, mechanical loading prediction based on the resultant ACC offers the important advantage of not being dependent on the accelerometer correct orientation, which is hard to assure when measuring daily physical activity [@Welk2005; @Wundersitz2015].

@@

To the best of our knowledge, only two studies have already attempted to develop pLR prediction models based on accelerometry data, and both studies only included walking activities. Accordingly to what we found, these studies also showed that the prediction error for these models is considerably larger than for the pGRF prediction. One of these researches investigated the the use of four different accelerometer placements to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], the other one found values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020]. Results from the models developed in both these investigations ranged from lower to higher prediction error compared to our results, which are approximately of 29.0\%. Although pLR is an important variable, known for a long time to affect the bone response to mechanical loading [@Bassey1998; @Turner1995], based on this and previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield satisfactory results. Nevertheless, past research have suggested that the bone response to pGRF and pLR is somewhat similar due to the high correlation between these variables [@Heikkinen2007]. Therefore, the pGRF prediction by accelerometry data, which is shown to be accurate, already incorporates some of the information about the effect of LR on the bone response to loading.

@@

Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, as shown in Table 2, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to distinguish the accelerometer data between walking and running, which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.

@@

In conclusion, pGRF could be accurately predicted by accelerometry-based regression models. These models are valid for the use of both the resultant vector and its vertical component, and using data from accelerometers placed at either the ankle, lower back or hip. The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications of monitoring the mechanical loading in several contexts and establishing the relationship of this loading and either a positive or negative osteoarticular adaptation. This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models were considerably lower than of the pGRF models.

>>>>

To assess the models prediction accuracy, mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE) were calculated. Although there is no standard threshold defining an acceptable error for mechanical loading prediction, based on previous studies we considered a MAPE <8\% to be an accurate prediction [@Fortune2014; @Neugebauer2014; @Ancillao2018]. Also, methods agreement was evaluated with Bland-Altman plots by plotting the difference between the actual (FP measured) and predicted values against their mean. Bias was computed as the mean of these differences and the limits of agreement were obtained using ±1.96 standard deviation of the mean between actual and predicted values [@Bland1986]. In order to test whether there was a consistent over or underestimation of the actual values, one-sample *t* tests were performed to assess if the difference between actual and predicted values are different than zero. Linear regressions were applied to identify if the magnitude of the mean between actual and predicted values influenced the magnitude of their difference. This was done to ascertain if the prediction accuracy was constant throughout the assayed magnitude range, which is known as proportional bias [@Giavarina2015]. These analyses were conducted separately for data from each accelerometer placement and outcome.

@@

Regression equations were developed to predict the mechanical loading variables based on accelerometry data. In the pGRF prediction equations, body mass and pACC were significant predictors and included in the models. Table 1 shows the regression coefficients, R^2^ and accuracy indices from models derived from accelerometers placed at ankle, lower back and hip, and from the resultant vector and its vertical component. In all pGRF prediction models, R^2^ values were `r max(GRF_accuracy$R2_cond)`, which shows that the predictor variables are able to explain the variability in the outcome variable to a great extent. Also, MAE ranged between `r myround(min(GRF_accuracy$MAE), 1)` N and `r myround(max(GRF_accuracy$MAE), 1)` N, and the MAPE from `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% to `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which is a smaller error than previously published equations [@Neugebauer2014; @Fortune2014], and similar to the error from equations previously published by our group, but which included walking only [@Veras2020]. RMSE were also computed and shown to range between `r myround(min(GRF_accuracy$RMSE), 1)` N and `r myround(max(GRF_accuracy$RMSE), 1)` N. Of all the pGRF predictions equations developed, the one based on hip-worn accelerometers and the resultant vector has shown the best performance, with an R^2^ of `r hip_GRF_accuracy$R2_cond`, and the lowest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).

@@

Also, to simplify the use of these equations and enhance their applicability, a software package was developed for the use in the R statistical software (REF). Included in this package are functionalities to read raw ACC data, process it and apply the equations to predict both pGRF and pLR. In the package website (ADDRESS) there are instructions on how to use it to obtain the predictions, namely in the 'Home' and 'Get started' sections.

@@

The purpose of this study was to develop pGRF and pLR prediction equations based on accelerometer data able to accurately estimate the mechanical loading during walking and running. The predictions were tested with accelerometers placed at different body locations and using both the resultant vector and its vertical component in a set of incremental speeds, ranging from slow walking to running. Our findings demonstrated that accelerometer-based regression models were a valid and accurate method to estimate pGRF during human locomotion, as all of the pGRF prediction models developed presented a MAPE below the 8\% threshold established by us. However, accelerometry-based equations developed to predict pLR did not yield trustworthy results.

@@

Our findings revealed that either ankle, lower back and hip were valid accelerometer placements to accurately predict pGRF. However, despite the similarities, hip was the accelerometer placement with the best prediction accuracy, which reinforces it as the preferable location to collect and analyse accelerometry-measured physical activity data. This placement is already well accepted for energy expenditure prediction and physical activity intensity classification [@Migueles2017] and has also shown a high mechanical loading prediction ability in previous studies [@Veras2020; @Neugebauer2018; @Neugebauer2012]. Besides the various placements, the current study also analysed predictions from both the resultant and vertical vectors. Although ACC and GRF from the vertical vector are the most studied and frequently associated with bone health [@Vainionpaa2006; @Heikkinen2007], the resultant vector of both ACC and GRF during locomotion is greatly determined by the vertical vector, and, as can be observed in Supplemental Figure S1, values from both vectors in this study were very similar. Also, mechanical loading prediction based on the resultant ACC offers the important advantage of not being dependent on the accelerometer correct orientation, which is hard to assure when measuring daily physical activity [@Welk2005; @Wundersitz2015].

@@

To the best of our knowledge, only two studies have already attempted to develop pLR prediction models based on accelerometry data, and both studies only included walking activities. Accordingly to what we found, these studies also showed that the prediction error for these models is considerably larger than for the pGRF prediction. One of these researches investigated the use of four different accelerometer placements (ankle, tibia, thigh and waist) to predict pLR and achieved a MAPE ranging from 23.4\% to 46.1\% [@Fortune2014], the other one found values from 19.5\% to 22.4\% using accelerometers at the lower back and hip [@Veras2020]. Results from the models developed in both these investigations ranged from lower to higher prediction error compared to our results, which are approximately of 29.0\%. Although pLR is an important variable, known for a long time to affect the bone response to mechanical loading [@Bassey1998; @Turner1995], based on this and previous studies [@Fortune2014; @Veras2020], its prediction through accelerometer data does not seem to yield trustworthy results. Nevertheless, past research has suggested that the bone response to pGRF and pLR is somewhat similar due to the high correlation between these variables [@Heikkinen2007]. Therefore, the pGRF prediction by accelerometry data, which is shown to be accurate, already incorporates some of the LR effect on the bone response to loading.

Results from our study offer some advantages compared to previous investigations: i) it includes walking and running, which are among the most usual types of physical activity performed daily; ii) all models developed, especially for pGRF prediction, presented a similar or better accuracy compared to the results from previous publications; iii) the accuracy of our models is constant throughout walking and running, and therefore they can be used adequately for both activities; iv) in order to apply our models there is no need to distinguish the accelerometer data between walking and running, which facilitates its use with daily living accelerometer data; and v) along with the creation of these models, a software was developed in the form of an R package to perform all the computations needed to get from the raw accelerometer data to the mechanical loading prediction.

@@

In conclusion, pGRF could be accurately predicted by accelerometry-based regression models. These models are valid for the use of both the resultant vector and its vertical component, and using data from accelerometers placed at either the ankle, lower back or hip. The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications of monitoring the mechanical loading in several contexts and establishing the relationship of this loading and either a positive or negative osteoarticular adaptation. This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models were unsatisfactory.
